---
title: "S&DS 363 HOMEWORK 4"
author: "George Hua, Valerie Nguyen, Michelle Liang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(MASS)
#library(biotools)
library(DiscriMiner) 
library(klaR)

crime <- read.csv("ohiocrimehm.csv", as.is=TRUE)
View(crime)
```
*response variables: 1=strongly disagree, 7= strongly agree*
*V10 (summer job provision for youths) V12 (recreation programs) V16 (treatment programs for addiction) V19 (job training and apprenticeship programs) V23 (help to troubled kids and families)*
*independent variables V64 (continuous) V67 (continuous) V86 (continuous) V87 (continuous) V70 (categorical) V71 (categorical) V72 (categorical) V73 (categorical)*

## #1
```{r}
#Make an object with specific variables
ohiocrime <- crime[,c(10,12,16,23,64,67,70,71,72,73,86,87)]
View(ohiocrime)
#Change columns from doubles to integers
ohiocrime <- sapply(ohiocrime, as.integer)
ohiocrime <- as.matrix(ohiocrime)
#Make response variable
responsevar <- as.numeric(ohiocrime[,1])
for (i in 2:5) {
  responsevar <- cbind(responsevar,as.numeric(ohiocrime[,i]))
}
#Make labels vector
responsevec <- c("Job provision", "Recreation program", "Treatment program", "Job training", "Help to troubled")

for (i in 1:5) {
  boxplot(ohiocrime[,i] ~ V87, data = ohiocrime, col = 'yellow', main = paste(responsevec[i], "By Income"), ylab = "")
  #calculate means using the tapply function - could also use the by function
  means <- tapply(ohiocrime[,i], ohiocrime[,9],mean)
  points(means, col = "red", pch = 19, cex = 1.2)
  text(x = c(1:4), y=means+.2, labels = round(means,2))
}

for (i in 1:5){
  boxplot(ohiocrime[,i] ~ V72, data = ohiocrime, col = 'yellow', main = paste(responsevec[i], "By Income"), ylab = "")
  #calculate means using the tapply function - could also use the by function
  means <- tapply(ohiocrime[,i], ohiocrime[,8],mean)
  points(means, col = "red", pch = 19, cex = 1.2)
  text(x = c(1:4), y=means+.2, labels = round(means,2))
}

interaction.plot(ohiocrime[,8], ohiocrime[,9], ohiocrime[,1])
```

## #2

```{r}
#Ran Two-way MANOVA for V71 and V72
man1 <- manova(as.matrix(ohiocrime[,c(9, 12)]) ~ ohiocrime[,"V71"] + ohiocrime[,"V72"] + ohiocrime[,"V71"]*ohiocrime[,"V72"])
#univariate result
summary.aov(man1)
```

